---
title: "Understanding intersectionality in community level predictors of Maternal Healthcare Utilization in India"
author: "Arnab K. Dey"
date: "3/2/2021"
output:
  slidy_presentation:
    highlight: tango
    slide_level: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r load-packages}
library(haven)
library(tidyverse)
library(summarytools)
library(compareGroups)
library(DiagrammeR)
library(party)
library(randomForest)
```

```{r load-data}
varlist_couples <- c("v001", "v020", "v021", "v024", "v312", "v025", "v012", "mv012", 
                     "v130", "v133", "mv133", "v205", "sm224", "s309",
                      "v202", "v203", "v204", "v205", "v191", "v602", "mv602", "v613", "mv613",
                     "v627", "mv627", "v628", "mv628", "v629", "mv629", "v632", "v627", 
                     "v628", "mv627", "mv628", "v190", "v106", "v701", "mv106", "s116")

df_master <- read_dta("C:/Users/arnab/OneDrive/Desktop/Analyses/Datasets/NFHS-4/Data/Couples Recode/IACR74DT/IACR74FL.dta", 
                      col_select = all_of(varlist_couples))
df_master <- as_factor(df_master)

```

```{r any modern method}

df_couples <- df_master

df_couples$any_modern <- with(df_couples, ifelse((v312 == "not using" | v312 == "rhythm/periodic abstinence" | 
                                               v312 == "withdrawal" | v312 == "lactational amenorrhea (lam)" | 
                                               v312 == "standard days"), 0, 1)) 

```

```{r num of children}
df_couples$num_child <- (df_couples$v202 + df_couples$v203 + df_couples$v204 + df_couples$v205)

```

```{r ideal number of children}
df_couples$w_ideal <- as.numeric(df_couples$v613)
df_couples$m_ideal <- as.numeric(df_couples$mv613)
```

```{r individual covariates}
## Age of women
df_couples$w_age <- df_couples$v012

## Age of Men
df_couples$m_age <- df_couples$mv012

# Education

## Education - Women
df_couples$w_edu <- as.numeric(df_couples$v133)

## Education - Men
df_couples$m_edu <- as.numeric(df_couples$mv133)

```


```{r couple level covariates}

## Religion
df_couples$islam <- ifelse(df_couples$v130 == "muslim", 1,0)

## Wealth
df_couples$wealth = ((df_couples$v191 - min(df_couples$v191))/((max(df_couples$v191)-min(df_couples$v191)))*100)

## Place of residence
df_couples <- df_couples %>% mutate(urban = ifelse(v025 == "urban", 1, 0))

## State 
df_couples <- df_couples %>% mutate(state = as.numeric(v024))

```

```{r dataset for analysis}

df_village <- df_couples %>% 
  select(v001, any_modern, num_child, w_ideal, m_ideal, m_age, w_age, w_edu, m_edu,
         islam, wealth, urban, state) %>% 
  group_by(v001) %>% 
  summarise_all(funs(mean))

df_village$v001 <- NULL

```

## Motivation

* Despite decades long efforts by the Government to push modern contraceptive methods, usage remains low in India
* Only 50% of couples use any modern method of contraceptive, Nationally
* The uptake of contraceptives varies significanlty by Socio-economic factors, access and Geography
* Huge heterogenieties exist is contraceptive use by region
* In this study, we aim to understand the factors that contribute to contraceptive use at a cluster level
* We also focus on understanding the intersectionalities of these factors as relates to contraceptive use

## Data Source

* Demographic Health Survey (DHS) Dataset for India - Couple's Dataset
* Data collected in 2015-16 
* Nationally representative sample of women in reproductive age (15-49 yrs)
* Who gave birth to a child in the last 5 years prior to the survey
* The Dataset comprises of 63,696 cases
* Clustered in 9,624 villages across the country
* Survey includes items asked to both partners (men and women) 

## Key Variables of Interest

* Outcome: Couple's use of any modern method of contraceptive
* Primary Independent Variable: Social Norms around fertility measured via men and women's ideal number of children (Continuous)
* **Key covariates**: 
+ Age of men and women (years)
+ Years of educational attainment of men and women 
+ Number of children that couples' have (continuous)
+ Religion (Islam / Other)
+ Household Wealth (Wealth index ranging from 0 to 100)
+ Place of Residence (Urban = 1 / Rural = 0)
+ State of residence (Categorical variable identifying one of the 36 states in India)

## Analytical Approach

* Undertake cluster-level analysis to understand factors that predict contraceptive use in villages
* Use Classification and Regression Tree (CART) analysis to understand interactions between predictors
* Undertake Random Forest analysis to understand feature importance in predicitng contraceptive use - at cluster level

## Methods

* **Step-1**: All variables where summarized at the cluster level by computing their mean at the village level
* **Step-2**: CART was undertaken with contraceptive use as the dependent variables
* **Step-3**: Random forests were generated with number of trees = 100, to get variable importance

# Key descriptive statistics

## Distribution of modern contraceptive uptake across clusters
```{r plot outcome}
ggplot(df_village, aes(x = any_modern)) +
  geom_histogram(binwidth = 0.08, fill = "#7a6374") + 
  xlab("Proportion of couples in the village who use any modern contraceptive") 
```

## Other covariates

```{r summary of variables}
tab1 <- compareGroups( ~ . , data = df_village)
createTable(tab1, digits = 2, show.n = FALSE)
```


## Findings from Classification and Regression Tree (CART) Analysis 

```{r generate-CART}
treemodel_mcpr = ctree(any_modern ~ . , data = df_village, controls=ctree_control(testtype = c("MonteCarlo"), 
                                                                  mincriterion = 0.95, savesplitstats = TRUE,
                                                                  maxdepth = 4))

```

Average proportion of modern contraceptive users at the village level is: `r mean(df_village$any_modern)`!

```{r plot CART, fig.width=20}

plot(treemodel_mcpr, type = "simple")

```


# Findings from the Random Forest Model

```{r random-forest}
set.seed(0112358)
forest_mcpr <- randomForest(any_modern~. , data = df_village, ntree = 100, importance = TRUE)
```

## Variable Importance Plot
```{r var importance plot}
varImpPlot(forest_mcpr)
```


## Variable importance coefficients
```{r Variable importance values}
importance(forest_mcpr)
```

## Implications

* A significant proportion of villages in India have very low proportions of modern contraceptive users 
* Our CART models suggest that social Norms around ideal number of children is an important predictor of contraceptive use at the village level
* These social norms also intersect with key socio-demographic variables such as age and number of children to influence contraceptive use
* Random Forest models suggests that household wealth could also be an important predictor of contraceptive use
* Findings underscore the need to engage with couples as well as communities to address social norms and goals around fertility

## Limitations

* CART is a highly unstable method and cut-points can vary substantially with changes in data
* We address this by using a Random Forest model - however that limits inference
* These methods can be computationally expensive as well
* Despite limitations, the broad findings from the study point have important implications 
* Finidngs help in understanding interesting intersections and call for deeper probes into the issue

# Questions?

# Thank You


