
### Problem Statement

Complications during delivery is a serious public health challenge that leads to Maternal and Neonatal 
Mortality. These complications include Breech Presentation, Prolonged Labor and Excessive Bleeding 
at the time of delivery. The problem is severe in Low and Middle Income countries like India where 
Maternal and Neonatal Mortality Rates are very high. 

A lot of Maternal and Neonatal deaths can be averted if Complications during delivery can be 
predicted correctly based on demographic data. Our goal in this project is to choose the right statistical learning
method(s) that would lead to the most accurate predictions 


### The Dataset

We use the Demographic Health Survey (DHS) Dataset for India which was released in 2015-16 and 
comprises of a nationally representative sample of women in the age group of 15-49 years who had a
birth in the last 5 years. The Dataset comprises of 699686 cases.

The dataset contains information on whether the women faced any complications related to 
Breech Presentation, Prolonged Labor and Excessive Bleeding during delivery. We recode these variables 
below to create a dichotomous variable that is coded 1 if women faced any of the three complications
and 0 if they did not face any of the complications during their last pregnancy


### Dependent Variable

The Dependent Variable will be created using the following variables: s439_1 (experience a breech presentation), 
s440_1 (experienced prolonged labor) and s441_1 (Experienced Excessive Bleeding). 


### Predictors

We consider the following predictors for the model:

Current age of the respondent in Yrs (v012)
Level of education (v133)
Household Wealth Category (s190s)

Total children ever born (v201)
Number of births in last 5 years (v208)
Age of respondent at first birth (v212)
Marriage to first birth interval in months (v220)
Ever had a terminated pregnancy (v228)

Weight of respondent in Kg (v437)
Height of respondent in Cm (v438)
Timing of first Antenatal Checkup in months (m13_1)
Number of antenatal visits during pregnancy (m14_1)
Place of delivery(m15_1)

Smokes cigarettes (v463a)
Smokes pipe (v463b)
Uses chewing tobacco (v463c)
Uses snuff (v463d)
Smokes cigars (v463e)
Gutkha/Paan masala with tobacco (v463f)
Paan with tobacco (v463g)
Drinks alcohol (s716)
Currently smoke bidis (s707)
Bidis smoked in last 24 hours (s708)
Type of tobacco used: Hookah (s710c)
Type of tobacco used: Khaini(s710e)

Experienced any less severe violence (D105A-C,J) by husband/partner (d106)
Experienced any severe violence (D105D-F) by husband/partner(d107)
Experienced any sexual violence (D105H-I,K) by husband/partner (d108)
Experienced any of listed husband/partner's actions (D110x series) (d111)
Husband/partner drinks alcohol (d113)
Pregnancy end in miscarriage, abortion, or stillbirth (s234)

```{r load-packages}
suppressMessages(library(haven))
suppressMessages(library(tidyverse))
suppressMessages(library(summarytools))
```

```{r variables of interest}
vars <- c("s439_1", "s440_1", "s441_1", "v012", "v133", "s190s", "v201", "v208", "v212",
          "v220", "v228", "v437", "v438", "m13_1", "m14_1", "m15_1", "v463a", "v463b",
          "v463c", "v463d", "v463e", "v463f", "v463g", "s716", "s707", "s708", "s710c", "s710e",
          "d106", "d107", "d108", "d111", "d113", "s234")
```




```{r load-dataset}
df <- read_dta("C:/Users/arnab/OneDrive/Desktop/Analyses/Datasets/NFHS-4/Data/Individual Recode/IAIR74DT/IAIR74FL.DTA", col_select = vars)
df <- haven::as_factor(df)

## Dropping cases for which the variable on complication in NA
df <- df[!(is.na(df$s439_1)),]
```



```{r Creating the Dependent Variable}
df$any_comp <- 0
df$any_comp[(df$s439_1 == "yes" | df$s440_1 == "yes" | df$s441_1 == "yes")] <- 1

df$any_comp <- factor(df$any_comp, levels = c(0,1), labels = c("No complication", "Any complication"))

freq(df$any_comp)
```



